# Cruncher and quality report {.unnumbered style="color:blue"}

## Overview

The cruncher is an additional "executable" module, not an R package. It can be launched via R, SAS.

Objective of the cruncher: 

- update a JDemetra+ workspace (with a given [revision policy](T-rev-policies-production.qmd))

- export the results (series and diagnostics),

without having to open the graphical interface and operate manually. Suitable for a production process.

## Installation procedure

-   Download the cruncher

Available here <https://github.com/jdemetra/jwsacruncher/releases>

Click on the zip code line of the latest release

-   Unzip locally (or on server)

## Help pages

Documentation is available here or click on the wiki icon on the Github page <https://github.com/jdemetra/jwsacruncher/wiki>

## Running the cruncher in R

Two R packages are currently available

-   rjwsacruncher on CRAN: workspace update and output production

-   Cruncher (https://github.com/InseeFr/JDCruncheR): same functions as rjwsacruncher      but adds a quality report

### Installation

-   rjwsacruncher 
```{r, eval = FALSE}
install.packages(rjwsacruncher)
```

-   JDCruncheR package: download the **.zip** or **.tar.gz** file from <https://github.com/InseeFr/JDCruncheR/releases>.

Additional packages needed

```{r, eval = FALSE}
install.packages(c("XLConnect", "XML"))
```

### Loading

```{r, eval = FALSE}
library(JDCruncheR)
```

or

```{r, eval = FALSE}
library(rjwsacruncher)
```

### Connecting the cruncher module

To connect the cruncher to the R package, the path to the bin directory containing the **cruncher.bat** file must be specified. This directory is available once the zip file has been unzipped.

```{r, eval = FALSE}
options(cruncher_bin_directory =
    "C:/Software/jwsacruncher-2.2.3/jdemetra-cli-2.2.3/bin")
```

-   checking the current value

```{r, eval = FALSE}
getOption("cruncher_bin_directory")
```

## Updating a workspace

The functions described in this section are identical for both packages.

### Running estimations

The general context 
- First estimation 

- Applying a [revision policy](T-rev-policies-production.qmd) to updated raw series

The function `cruncher_and_param()` allows to do that

```{r, eval = FALSE}

cruncher_and_param(workspace = "D:/my_folder/my_ws.xml",
                   rename_multi_documents = FALSE,
                   policy = "lastoutliers" #name of the revision policy
                   log= my_log_file.txt)
```

To use the documentation, compute `help()` or `?function`:

```{r, eval = FALSE}
?cruncher_and_param
help(cruncher_and_param)
```

Before running SA estimations, set the export options.

### Configuring output options

After updating the workspace with the selected revision policies, the cruncher generates output 
- series (csv files) 
- diagnostics and parameters (demetra_m.csv file)

These files will be created in the workspace's repository
path="My_Workspace\Output\SAProcessing"


#### Selecting time series to export

```{r, eval = FALSE}
# returns names of the currently exported series 
getOption("default_tsmatrix_series")
# example of setting this option
options(default_tsmatrix_series = c("sa", "sa_f"))
# only seasonally adjsuted series ("sa") and its forecasts ("sa_f") will be exported

```

#### Selecting diagnostics and parameters to exoprt

```{r, eval = FALSE}
# returns names of the currently exported diagnostics and parameters
getOption("default_matrix_item")
# example of setting this option
options(default_matrix_item = c("likelihood.aic",
                                "likelihood.aicc",
                                "likelihood.bic",
                                "likelihood.bicc"))

```
=

## Quality report with JDCruncheR

The JDCruncheR package also:

-   computes a quality score

-   creates a quality report from the diagnostics produced by JDemetra+

### Main steps

The three main functions of the package are:

-   `extract_QR` to extract the quality report from the csv file (demetra_m.csv) that contains all JD+ diagnostics;

-   `compute_score` to compute a weighted score based on the diagnostics

-   `export_xlsx` to export the quality report.

```{r, eval = FALSE}
# choose the demetra_m.csv file generated by the cruncher
QR <- extract_QR()
QR

?compute_score # to see how the score is calculated (formula)
QR <- compute_score(QR,
                    n_contrib_score = 3)

QR

QR <- sort(QR, decreasing = TRUE, sort_variables = "score")
export_xlsx(QR,
            file_name = "U:/quality_report.xls")
```

### Piling up results for several SAP's

When working with several workspaces (or SAPs), quality reports can be piled up with the function `rbind()` or by creating a mQR_matrix object with the function `mQR_matrix()`

```{r, eval = FALSE}
QR1 <- extract_QR()
QR2 <- extract_QR()
mQR <- mQR_matrix(QR1, QR2)
mQR
# naming each object
names(mQR) <- c("report_1", "report_2")
# Equivalent to: 
mQR <- mQR_matrix(report_1 = QR1, report_2 = QR2)
mQR

# score calculation for all reports
mQR <- compute_score(mQR,
                    n_contrib_score = 3)
export_xlsx(mQR,
            export_dir = "U:/")
```

### Conditionnal score

Missing values can be ignored and conditions can be set for indicators:

```{r, eval = FALSE}
# oos_mse weight reduced to 1 when the other 
# indicators are "Bad" ou "Severe"
condition1 <- list(indicator = "oos_mse",
                   conditions = c("residuals_independency",
                                  "residuals_homoskedasticity",
                                  "residuals_normality"),
                   conditions_modalities = c("Bad","Severe"))
BQ <- compute_score(BQ, n_contrib_score = 5,
                    conditional_indicator = list(condition1),
                    na.rm = TRUE)
```

### Customize the score computation

Practical steps if you want to customize the score computation (see package documentation in R)

-   select your indicators of interest

-   adjust "good", "bad"...threshold in JD+ GUI if necessary

-   by default good=0, uncertain=1, bad or severe=3

-   change this grading system and/or the weights directly in the package functions

-   rebuild your package


## List of exportable diagnostics and parameters

```{r, eval=FALSE}

options(default_matrix_item = c("period", "span.start", "span.end", "span.n", "span.missing",
                                "espan.start", "espan.end", "espan.n", "log", "adjust", "regression.lp",
                                "regression.ntd", "regression.nmh", "regression.td-derived",
                                "regression.td-ftest", "regression.easter", "regression.nout",
                                "regression.noutao", "regression.noutls", "regression.nouttc",
                                "regression.noutso", "regression.td(*):4", "regression.out(*)",
                                "regression.user(*)", "likelihood.neffectiveobs", "likelihood.np",
                                "likelihood.logvalue", "likelihood.adjustedlogvalue", "likelihood.ssqerr",
                                "likelihood.aic", "likelihood.aicc", "likelihood.bic", "likelihood.bicc",
                                "residuals.ser", "residuals.ser-ml", "residuals.mean", "residuals.skewness:3",
                                "residuals.kurtosis:3", "residuals.dh", "residuals.lb", "residuals.lb2:3",
                                "residuals.seaslb", "residuals.bp", "residuals.bp2", "residuals.seasbp",
                                "residuals.nudruns", "residuals.ludruns", "residuals.nruns",
                                "residuals.lruns", "arima", "arima.mean", "arima.p", "arima.d",
                                "arima.q", "arima.bp", "arima.bd", "arima.bq", "arima.phi(*)",
                                "arima.bphi(*)", "arima.th(*)", "arima.bth(*)", "decomposition.seasonality",
                                "decomposition.parameters_cutoff", "decomposition.model_changed",
                                "decomposition.tvar-estimator", "decomposition.tvar-estimate",
                                "decomposition.tvar-pvalue", "decomposition.savar-estimator",
                                "decomposition.savar-estimate", "decomposition.savar-pvalue",
                                "decomposition.svar-estimator", "decomposition.svar-estimate",
                                "decomposition.svar-pvalue", "decomposition.ivar-estimator",
                                "decomposition.ivar-estimate", "decomposition.ivar-pvalue", "decomposition.tscorr-estimator",
                                "decomposition.tscorr-estimate", "decomposition.tscorr-pvalue",
                                "decomposition.ticorr-estimator", "decomposition.ticorr-estimate",
                                "decomposition.ticorr-pvalue", "decomposition.sicorr-estimator",
                                "decomposition.sicorr-estimate", "decomposition.sicorr-pvalue",
                                "decomposition.ar_root(*)", "decomposition.ma_root(*)", "method",
                                "variancedecomposition.cycle", "variancedecomposition.seasonality",
                                "variancedecomposition.irregular", "variancedecomposition.tdh",
                                "variancedecomposition.others", "variancedecomposition.total",
                                "diagnostics.logstat", "diagnostics.levelstat", "diagnostics.fcast-insample-mean",
                                "diagnostics.fcast-outsample-mean", "diagnostics.fcast-outsample-variance",
                                "diagnostics.seas-lin-f", "diagnostics.seas-lin-qs", "diagnostics.seas-lin-kw",
                                "diagnostics.seas-lin-friedman", "diagnostics.seas-lin-periodogram",
                                "diagnostics.seas-lin-spectralpeaks", "diagnostics.seas-si-combined",
                                "diagnostics.seas-si-evolutive", "diagnostics.seas-si-stable",
                                "diagnostics.seas-res-f", "diagnostics.seas-res-qs", "diagnostics.seas-res-kw",
                                "diagnostics.seas-res-friedman", "diagnostics.seas-res-periodogram",
                                "diagnostics.seas-res-spectralpeaks", "diagnostics.seas-res-combined",
                                "diagnostics.seas-res-combined3", "diagnostics.seas-res-evolutive",
                                "diagnostics.seas-res-stable", "diagnostics.seas-i-f", "diagnostics.seas-i-qs",
                                "diagnostics.seas-i-kw", "diagnostics.seas-i-periodogram", "diagnostics.seas-i-spectralpeaks",
                                "diagnostics.seas-i-combined", "diagnostics.seas-i-combined3",
                                "diagnostics.seas-i-evolutive", "diagnostics.seas-i-stable",
                                "diagnostics.seas-sa-f", "diagnostics.seas-sa-qs", "diagnostics.seas-sa-kw",
                                "diagnostics.seas-sa-friedman", "diagnostics.seas-sa-periodogram",
                                "diagnostics.seas-sa-spectralpeaks", "diagnostics.seas-sa-combined",
                                "diagnostics.seas-sa-combined3", "diagnostics.seas-sa-evolutive",
                                "diagnostics.seas-sa-stable", "diagnostics.seas-sa-ac1", "diagnostics.td-sa-all",
                                "diagnostics.td-sa-last", "diagnostics.td-i-all", "diagnostics.td-i-last",
                                "diagnostics.td-res-all", "diagnostics.td-res-last", "diagnostics.ic-ratio-henderson",
                                "diagnostics.ic-ratio", "diagnostics.msr-global", "diagnostics.msr(*)",
                                "decomposition.trendfilter", "decomposition.seasfilter", "m-statistics.m1",
                                "m-statistics.m2", "m-statistics.m3", "m-statistics.m4", "m-statistics.m5",
                                "m-statistics.m6", "m-statistics.m7", "m-statistics.m8", "m-statistics.m9",
                                "m-statistics.m10", "m-statistics.m11", "m-statistics.q", "m-statistics.q-m2",
                                "diagnostics.basic checks.definition:2", "diagnostics.basic checks.annual totals:2",
                                "diagnostics.visual spectral analysis.spectral seas peaks", "diagnostics.visual spectral analysis.spectral td peaks",
                                "diagnostics.regarima residuals.normality:2", "diagnostics.regarima residuals.independence:2",
                                "diagnostics.regarima residuals.spectral td peaks:2", "diagnostics.regarima residuals.spectral seas peaks:2",
                                "diagnostics.outliers.number of outliers:2", "diagnostics.out-of-sample.mean:2",
                                "diagnostics.out-of-sample.mse:2", "diagnostics.m-statistics.q:2",
                                "diagnostics.m-statistics.q-m2:2", "diagnostics.seats.seas variance:2",
                                "diagnostics.seats.irregular variance:2", "diagnostics.seats.seas/irr cross-correlation:2",
                                "diagnostics.residual seasonality tests.qs test on sa:2", "diagnostics.residual seasonality tests.qs test on i:2",
                                "diagnostics.residual seasonality tests.f-test on sa (seasonal dummies):2",
                                "diagnostics.residual seasonality tests.f-test on i (seasonal dummies):2",
                                "diagnostics.combined seasonality test.combined seasonality test on sa:2",
                                "diagnostics.combined seasonality test.combined seasonality test on sa (last 3 years):2",
                                "diagnostics.combined seasonality test.combined seasonality test on irregular:2",
                                "diagnostics.residual trading days tests.f-test on sa (td):2",
                                "diagnostics.residual trading days tests.f-test on i (td):2",
                                "diagnostics.quality"
))

```


